The Role of Shallow Processing in Agreement Attraction

### 2020-01-15 agremeent_draft_dec17/draft_dec17.Rnw
minor fix

<<>>=
# load packages
library(patchSynctex)
library(knitcitations)
library(car, warn.conflicts = FALSE)
library(MASS)
library(brms)
library(xtable)
library(ggpubr)
library(languageR)
library(tidyverse)
library(gdata)
library(magrittr)
library(ggplot2)
theme_set(theme_bw())

@@

<<loadExperiment2Data>>=
fname_data2 <- "../workspace_exp2/exp_data.rds"
data_exp2 <- readRDS(file = fname_data2)
fname_form <- "../workspace_exp2/exp_form.rds"
form_exp2 <- readRDS(file = fname_form)

#! No idea what this is about. Excluding for now.
# form_exp2 <- form_exp2[-c(56),]

@@

<<badSubjects2>>=
# identify participants 
bad_subjects2 <- subset(avg_by_subj_wide2, delta_dc <= 0.25 ) %>% .$subject

data_exp2_clean <- data_exp2 %>% subset(!subject %in% bad_subjects)

@@

df_merge_exp2 <- data_exp2_clean %>% ungroup() %>% 
                      dplyr::select(source=experiment, grammatical, attractor_num, # condition,
                                    subject, item=Item,
                                    ResponseYes, RT)
df_merge_exp2$experiment <- "Experiment 2"
df_merge_exp2$grammatical <- with(df_merge_exp2, ifelse(grammatical == "gram", 
                                                        "grammatical",
                                                        "ungrammatical"))
df_merge_exp2$attractor_num <- with(df_merge_exp2, ifelse(attractor_num == "pl", 
                                                          "plural", 
                                                          "singular"))
df_merge_exp2$item %<>% as.integer()
df_merge_exp2$subject %<>% as.character()

@@

df_merged2 <- dplyr::bind_rows(df_merge_exp1, df_merge_exp2)
df_merged2$subject %<>% as.factor()
df_merged2$item %<>% as.factor()

df_exp2_na_nofillers <- subset(df_merged2, is.na(ResponseYes) & source != "filler")

df_merged2_nonna <- df_merged2 %>% subset(!is.na(ResponseYes))

@@

Eighty native speakers of Turkish in Experiment 1. The mean age is \Sexpr{mean(as.numeric(as.character(form_exp2$Age)))}, ranging from \Sexpr{min(as.numeric(as.character(form_exp2$Age)))} to \Sexpr{max(as.numeric(as.character(form_exp2$Age)))}. Across participants, a total \Sexpr{sum(is.na(df_exp2_na_nofillers$ResponseYes))} items were not responded to before the 5 second deadline by \Sexpr{length(unique(df_exp2_na_nofillers$subject))} different participants. These items were not included in the analysis. Experiments in this paper were carried out following the Declaration of Helsinki and the regulations concerning ethics in research in Bo\u{g}azi\c{c}i University. All participants provided informed consent. They were tested online, on IbexFarm, using their preferred platform, % Find citation for using IbexFarm and online testing etc?

@@

\ex. \label{item:exp2ExperimentalItems}
%
\a. \textit{Plural Attractor, Plural Verb}\label{item:exp2expitem-plpl}\\ 
  \gll *Tut-tuk-lar-ı aşcı mutfak-ta sürekli zıpl-ıyor-lar.\\ 
  hire-\textsc{nmlz}-\textsc{pl}-\textsc{poss}  cook kitchen-\textsc{loc} non-stop  jump-\textsc{prog}-\textsc{pl}.\\
  \glt \textit{`The cook that they hired were jumping in the kitchen non-stop.'}
%
\b. \textit{Plural Attractor, Singular Verb}\label{item:exp2expitem-plsg}\\ 
  \gll Tut-tuk-lar-ı aşcı mutfak-ta sürekli zıpl-ıyor.\\ 
  hire-\textsc{nmlz}-\textsc{pl}-\textsc{poss}  cook kitchen-\textsc{loc} non-stop  jump-\textsc{prog}.\\
  \glt \textit{`The cook that they hired was jumping in the kitchen non-stop.'}
%
\c. \textit{Singular Attractor, Plural Verb}\label{item:exp2expitem-sgpl}\\
  \gll *Tut-tuğ-u aşcı mutfak-ta sürekli zıpl-ıyor-lar.\\ 
  hıre-\textsc{nmlz}-\textsc{poss}  cook kitchen-\textsc{loc} non-stop  jump-\textsc{prog}-\textsc{pl}.\\
  \glt \textit{`The cook that they hired were jumping in the kitchen non-stop.'}
%
\d. \textit{Singular Attractor, Singular Verb}\label{item:exp2expitem-sgsg}\\ 
  \gll Tut-tuğ-u aşcı mutfak-ta sürekli zıpl-ıyor.\\ 
  hıre-\textsc{nmlz}-\textsc{poss}  cook kitchen-\textsc{loc} non-stop  jump-\textsc{prog}.\\
  \glt \textit{`The cook that they hired was jumping in the kitchen non-stop.'}

@@

In Experiment 2, the same procedure as Experiment 1 has been followed with regards to how sentences are presented within the experiment.

@@

@@

# ## test model parameterization using a simple GLM first
# m <- glm(ResponseYes ~ cEndsInConsonant * cUngrammatical * cAttractorPlural,
#          data = df_merged_nofillers,
#          family = binomial("probit"))
# summary(m)

@@

# TODO: Data seems to contain NAs in relevant columns; fix it
m_responses2 <- brm(ResponseYes ~ cVerbalAttractor * cUngrammatical * cAttractorPlural + 
                                 (cUngrammatical * cAttractorPlural + 1| subject) +
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers2,
                   family = bernoulli("probit"),
                   file = fname_exp2_responses, 
                   chains = n_chains, cores = n_cores, iter = n_iter)

@@

    p_m_response_exp2 <- 
      create_model_coefs_plot( m_responses2, 
            plot_stats = T, map_names = contrast_names2_le,
            expand_right = 4.5, expand_top = 1.5, x_stat_adjust = .75,
            x_breaks = c(-3,-2,-1, 0, 1) ) + 
            xlab("Estimate (log-odds)")

>>>>

# load packages
library(patchSynctex)
library(knitcitations)
library(car, warn.conflicts = FALSE)
library(MASS)
library(brms)
library(xtable)
library(ggpubr)
library(languageR)
library(tidyverse)
library(gdata)
library(magrittr)
library(ggplot2)
theme_set(theme_bw())

@@

<<loadExperiment2Data>>=

fname_data2 <- "../workspace_exp2/exp_data.rds"
data_exp2 <- readRDS(file = fname_data2)
fname_form <- "../workspace_exp2/exp_form.rds"
form_exp2 <- readRDS(file = fname_form)

@@

<<badSubjects2>>=

# identify bad participants 
bad_subjects2 <- subset(avg_by_subj_wide2, delta_dc <= 0.25 ) %>% .$subject

data_exp2_clean <- data_exp2 %>% subset(!subject %in% bad_subjects2)

@@

df_merge_exp2 <- data_exp2_clean %>% ungroup() %>% 
                      dplyr::select(source=experiment, grammatical, attractor_num, # condition,
                                    subject, item=Item,
                                    ResponseYes, RT)
df_merge_exp2$experiment <- "Experiment 2"
df_merge_exp2$grammatical %<>% dplyr::recode("gram"="grammatical", "ungram"="ungrammatical")
df_merge_exp2$attractor_num %<>% dplyr::recode("pl"="plural", "sg"="singular")
df_merge_exp2 %<>% mutate( item = sprintf("I2[%s]", as.character(item)), subject = as.character(subject) )

@@

df_merge_exp1$item %<>% as.character()
df_merged2 <- dplyr::bind_rows(df_merge_exp1, df_merge_exp2)
df_merged2 %<>% mutate( subject = as.character(subject), item = as.character(item) )

df_exp2_na_nofillers <- df_merged2 %>% subset(is.na(ResponseYes) & source != "filler")
df_merged2_nonna <- df_merged2 %>% subset(!is.na(ResponseYes))

@@

We recruited 79 native speakers of Turkish to participate in Experiment 2. Their average age was \Sexpr{round(mean(form_exp2$Age))}, ranging from \Sexpr{min(form_exp2$Age)} to \Sexpr{max(form_exp2$Age)}.

Experiments in this paper were carried out following the Declaration of Helsinki and the regulations concerning ethics in research in Bo\u{g}azi\c{c}i University. All participants provided informed consent. They were tested online, on IbexFarm, using their preferred platform, % Find citation for using IbexFarm and online testing etc?

@@

\ex. \label{item:exp2ExperimentalItems}
%
\a. \textsc{Plural Attractor, Ungrammatical (Plural Verb)}\label{item:exp2expitem-plpl}\\ 
  \gll *Tut-tuk-lar-ı aşcı mutfak-ta sürekli zıpl-ıyor-lar.\\ 
  hire-\textsc{nmlz}-\textsc{pl}-\textsc{poss}  cook kitchen-\textsc{loc} non-stop  jump-\textsc{prog}-\textsc{pl}.\\
  \glt \textit{`The cook that they hired were jumping in the kitchen non-stop.'}
%
\b. \textit{Plural Attractor, Grammatical (Singular Verb)}\label{item:exp2expitem-plsg}\\ 
  \gll Tut-tuk-lar-ı aşcı mutfak-ta sürekli zıpl-ıyor.\\ 
  hire-\textsc{nmlz}-\textsc{pl}-\textsc{poss}  cook kitchen-\textsc{loc} non-stop  jump-\textsc{prog}.\\
  \glt \textit{`The cook that they hired was jumping in the kitchen non-stop.'}
%
\c. \textit{Singular Attractor, Ungrammatical (Plural Verb)}\label{item:exp2expitem-sgpl}\\
  \gll *Tut-tuğ-u aşcı mutfak-ta sürekli zıpl-ıyor-lar.\\ 
  hıre-\textsc{nmlz}-\textsc{poss}  cook kitchen-\textsc{loc} non-stop  jump-\textsc{prog}-\textsc{pl}.\\
  \glt \textit{`The cook that they hired were jumping in the kitchen non-stop.'}
%
\d. \textit{Singular Attractor, Grammatical (Singular Verb)}\label{item:exp2expitem-sgsg}\\ 
  \gll Tut-tuğ-u aşcı mutfak-ta sürekli zıpl-ıyor.\\ 
  hıre-\textsc{nmlz}-\textsc{poss}  cook kitchen-\textsc{loc} non-stop  jump-\textsc{prog}.\\
  \glt \textit{`The cook that they hired was jumping in the kitchen non-stop.'}

@@

% to-do: some more details.
% In Experiment 2, the same procedure as Experiment 1 has been followed with regards to how sentences are presented within the experiment.

@@

Across participants, a total \Sexpr{sum(is.na(df_exp2_na_nofillers$ResponseYes))} items were not responded to before the 5 second deadline by \Sexpr{length(unique(df_exp2_na_nofillers$subject))} different participants. These items were not included in the analysis.

@@

# ## test model parameterization using a simple GLM first
# m <- glm(ResponseYes ~ cVerbalAttractor * cUngrammatical * cAttractorPlural,
#          data = df_merged_nofillers2,
#          family = binomial("probit"))
# summary(m)

@@

# TODO: Data seems to contain NAs in relevant columns; fix it
m_responses2 <- brm(ResponseYes ~ cVerbalAttractor * cUngrammatical * cAttractorPlural + 
                                 (cUngrammatical * cAttractorPlural + 1| subject) +
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers2 %>% subset(!is.na(ResponseYes)),
                   family = bernoulli("probit"),
                   file = fname_exp2_responses, 
                   chains = n_chains, cores = n_cores, iter = n_iter)

@@

p_m_response_exp2 <- 
  create_model_coefs_plot( m_responses2, 
        plot_stats = T, map_names = contrast_names2_le,
        expand_right = 4.5, expand_top = 1.5, x_stat_adjust = .75,
        x_breaks = c(-3,-2,-1, 0, 1) ) + 
        xlab("Estimate (probit)")

### 2020-01-15 agremeent_draft_dec17/draft_dec17.tex
minor fix

@@

Eighty native speakers of Turkish in Experiment 1. The mean age is 21.22, ranging from 0 to 31. Across participants, a total 55 items were not responded to before the 5 second deadline by 41 different participants. These items were not included in the analysis. Experiments in this paper were carried out following the Declaration of Helsinki and the regulations concerning ethics in research in Bo\u{g}azi\c{c}i University. All participants provided informed consent. They were tested online, on IbexFarm, using their preferred platform, % Find citation for using IbexFarm and online testing etc?

@@

@@

\autoref{fig:exp2AvgResponse} shows the average proportions of 'acceptable' responses by experimental conditions for both Experiment 1 and Experiment 2. We see that there is no effect resembling agreement attraction in the ungrammatical conditions of Experiment 2. Ungrammatical conditions showed comparable effects (0.07 and 0.07 for plural and singular attractors respeectively) regardless of the attractor's number.

>>>>

% to-do:
% Importantly, even if there was an interaction with experiment, it's relatively small, so it wouldn't reduce the other interaction to 0.

@@

We recruited 79 native speakers of Turkish to participate in Experiment 2. Their average age was 21, ranging from 18 to 31.

Experiments in this paper were carried out following the Declaration of Helsinki and the regulations concerning ethics in research in Bo\u{g}azi\c{c}i University. All participants provided informed consent. They were tested online, on IbexFarm, using their preferred platform, % Find citation for using IbexFarm and online testing etc?

@@

Across participants, a total 55 items were not responded to before the 5 second deadline by 45 different participants. These items were not included in the analysis.

@@

\autoref{fig:exp2AvgResponse} shows the average proportions of 'acceptable' responses by experimental conditions for both Experiment 1 and Experiment 2. We see that there is no effect resembling agreement attraction in the ungrammatical conditions of Experiment 2. Ungrammatical conditions showed comparable effects (0.05 and 0.06 for plural and singular attractors respeectively) regardless of the attractor's number.
